New lower bounds for Schur and weak Schur numbers

### 2021-05-07 article/article.tex
notations
relecture (pas encore terminée partie weak schur)

\usepackage[english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

\newcommand{\WS}{\mathit{WS}}

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
apply Rowley's construction and to get explicit partitions improving lower bounds as well as the growth rate for 
both Schur numbers and Ramsey numbers \(R_n(3)\). We also developed a new method to improve lower bounds on 
weak Schur numbers. Furthermore, this paper tries to analyze former works on the subject based on the principle 
that good partitions into \(n+1\) subsets start with good partitions into \(n\) subsets. We show that exceeding the 
previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption upon imposing certain conditions 
on the good 5-subsets partition. The new lower bounds include \(S(9) \geqslant 17803\), \(S(10) \geqslant 60948\),
\(\WS (9) \geqslant 22536 \) and \(\WS (10) \geqslant 71214 \).

@@

\include{definitions_notations.tex}

@@

\qquad These new results come from an extension of Rowley's template-based approach for Ramsey graphs and 
Schur numbers which is relatively new. Therefore, we would not be surprised if lower bounds are later improved 
using better templates. Moreover, studying specifically \(S^+(n)\) and \(\WS^+(n)\) might be of interest as they
 are closely related to Schur and weak Schur numbers. In order to find new templates, algorithms based on 
randomness such as Monte-Carlo algorithms may prove to be very useful. This could be the subject of a future work.

@@

\bibliography{biblio}
\bibliographystyle{ieeetr}

>>>>

\usepackage[english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

\newcommand{\WS}{\mathit{WS}}
\newcommand{\docarraystretch}{1.2}
\newcommand{\resetarraystretch}{\renewcommand{\arraystretch}{\docarraystretch}}
\resetarraystretch

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
get explicit partitions improving lower bounds as well as the growth rate for both Schur numbers and Ramsey 
numbers \(R_n(3)\). We also developed a new method to improve lower bounds for weak Schur numbers. 
Furthermore, this paper tries to analyze former works on weak Schur numbers based on the principle that good 
partitions into \(n+1\) weakly sum-free subsets start with a good partition into \(n\) weakly sum-free subsets. 
We show that exceeding the previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption 
upon imposing certain conditions on the good 5-subsets partitions. The new lower bounds include 
\(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536 \) and 
\(\WS (10) \geqslant 71\,214 \).

@@

% \include{definitions_notations.tex}

@@

\qquad These new results come from an extension of Rowley's template-based approach for Ramsey graphs and 
Schur numbers which is relatively new. Therefore, we will not be surprised if lower bounds are later improved 
using better templates. Moreover, studying specifically \(S^+(n)\) and \(\WS^+(n)\) might be of interest as they
 are closely related to Schur and weak Schur numbers. In order to find new templates, algorithms based on 
randomness such as Monte-Carlo algorithms may prove to be very useful. This could be the subject of a future work.

@@

\bibliography{biblio}
\bibliographystyle{IEEEtran}

### 2021-05-07 article/introduction.tex
relecture (pas encore terminée partie weak schur)

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) in \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
This article focuses on the lower bounds for these numbers.

@@

\par
As for smaller numbers, the best lower bounds were obtained by conducting a computer search. Eliahou \cite{ELIAHOU2012175}, 
Rafilipojaona \cite{Rafilipojaona} and Bouzy \cite{Bouzy2015AnAP} improved lower bounds with Monte-Carlo methods. It 
has been the main approch in the previous decade. 
This search for weakly sum-free partitions relied on the recursive assumption that a good weakly sum-free partition into \(n+1\) 
colors starts with a good weakly sum-free partition into \(n\) colors. This assumption was necessary in order for the Monte-Carlo
approch to yield results. However, we show in the \hyperlink{sat}{last section} how this search space may not countain
the optimal solution. Thus further work using similar methods might need to put that assumption aside and 
find a different search space.

\par
In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers which generalizes Abbott and 
Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. Rowley also gives two 
inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements over previous lower 
bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{ELIAHOU2012175} & 
    \cite{ELIAHOU2012175} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}

@@

\par
The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
This gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a 
special case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
We first explain Rowley's template-based construction in the context of Schur numbers and then give 
new templates, thus providing new lowers bounds and inequalities as well as showing that the growth rates 
for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
We then  generalize the concept of templates to weak Schur numbers and provide new lower bounds for weak 
Schur numbers.

\par
Finally, we analyze the significant difference between new lower bounds obtained with templates and the former 
lower bounds obtained by computer search and we provide evidence which indicate that the main assumption 
made in those articles removes the optimal partitions from the search space.

@@

We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers.

\begin{definition}
A subset \(A\) of \(\mathbb{N}\) is said to be \textit{sum-free} when:
\[ \forall (a,b) \in A^2 \text{, } a+b \notin A\]
\end{definition}

\begin{definition}
A subset \(B\) of \(\mathbb{N}\) is said to be \textit{weakly sum-free} when:
\[ \forall (a,b) \in B^2 \text{, } a \neq b \Longrightarrow a+b \notin B\]
\end{definition}

@@

\par
Schur proved in \cite{Schur1917} that given a number of subsets \(n\), there exists a value of \(p\)
such that there exists no partition of \([\![1,q]\!]\) into \(n\) sum-free subsets for any \(q \geqslant p\). A similar
property holds for weakly sum-free subsets (reference necessaire). These observations lead to the following definitions.

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a greatest integer that we denote \(S(n)\) (\textit{resp. \(\WS (n)\)}) such that
\([\![1,S(n)]\!]\) (resp. \([\![1, \WS (n)]\!]\)) can be partitioned into \(n\) sum-free subsets (resp. weakly sum-free
subsets). \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number} and \textit{\(\WS (n)\) the
\(n\)\textsuperscript{th} weak
Schur number}.
\end{definition}

Given a partition of \([\![1, p]\!]\) in \(n\) subsets, we generally denote these subsets \(A_1, ..., A_n\). We also denote
\(m_i = \min(A_i)\). By ordering the subsets, we mean assuming that \(m_1 < ... < m_n\). However, if not specified we do 
not make this hypothesis since we do not always consider partitions in which every subset plays a symmetric role.

>>>>

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
This article focuses on the lower bounds for these numbers.

@@

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade. This search for weakly 
sum-free partitions relied on the recursive assumption that a good weakly sum-free partition into \(n+1\) colors 
starts with a good weakly sum-free partition into \(n\) colors. 
However, we show in the \hyperref[SearchSpace]{last section} that this search space may not contain the optimal 
solutions. Thus further works using similar methods might need to put that assumption aside and find a different 
search space.

\par
In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. 
Rowley also gives two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}

@@

\qquad The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
It gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special 
case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
In the \hyperref[Schur]{first section}, we explain Rowley's template-based construction in the context of 
Schur numbers and then give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{second section}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

\par
Finally, in the \hyperref[SearchSpace]{third section}, we analyze the significant difference between new lower 
bounds obtained with templates and the former lower bounds obtained by computer search and we provide 
evidence which indicate that the main assumption made in those articles removes the optimal partitions from the 
search space.

@@

\qquad We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers.

\begin{definition}
A subset \(A\) of \(\mathbb{N}\) is said to be \textit{sum-free} if
\[ \forall (a,b) \in A^2 \text{, } a+b \notin A\]
\end{definition}

\begin{definition}
A subset \(B\) of \(\mathbb{N}\) is said to be \textit{weakly sum-free} if
\[ \forall (a,b) \in B^2 \text{, } a \neq b \Longrightarrow a+b \notin B\]
\end{definition}

@@

\par
Schur proved in \cite{Schur1917} that given a number of subsets \(n\), there exists a value of \(p\)
such that there exists no partition of \([\![1,q]\!]\) into \(n\) sum-free subsets for any \(q \geqslant p\). A similar
property holds for weakly sum-free subsets \cite{Irving1973}. These observations lead to the following definitions.

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a greatest integer denoted by \(S(n)\) (\textit{resp. \(\WS (n)\)}) such that
\([\![1,S(n)]\!]\) (resp. \([\![1, \WS (n)]\!]\)) can be partitioned into \(n\) sum-free subsets (resp. weakly sum-free
subsets). \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number} and \textit{\(\WS (n)\) the
\(n\)\textsuperscript{th} weak Schur number}.
\end{definition}

Given a partition of \([\![1, p]\!]\) into \(n\) subsets, we generally denote these subsets \(A_1, ..., A_n\). We also denote
\(m_i = \min(A_i)\). By ordering the subsets, we mean assuming that \(m_1 < ... < m_n\). However, if not specified we do 
not make this hypothesis since we do not always consider partitions in which every subset plays a symmetric role.

### 2021-05-07 article/schur.tex
relecture (pas encore terminée partie weak schur)

\begin{definition}
A SF-template with width \(p\) and \(n\) colors is defined as a partition of \( [\![1,p]\!]\) into \(n\) sum-free subsets 
\(A_1, A_2, ..., A_n\) verifying:
\[
\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
\Longrightarrow x+y-p \notin A_i
\]
\end{definition}

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty[\![\). We define \(S^+(n)\) as the maximal width of a SF-template with \(n\) colors. 
	\(S^+(n)\) is well defined and verifies:
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with width \(p\) and \(k+1\) colors 
	and a partition of \([\![1,q]\!]\) into \(n\) sum-free subsets then there exists a partition of \([\![1,pq+m_{k+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{k+1}\) is the minimum of the special subset in the \(SF\)-template.
\end{theorem}

@@

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They vertically 
extend a sum-free partition by repeating it and they use an other sum-free partition to color the other half according 
to the line number. This way the "blocks" act as safe areas for each other. We give here an example for \(p = 9\), 
\(q = 4\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) (2 S(2) + 1) + S(2)\), both with Abbott 
and Hanson's construction and with a SF-template which is not of this type. In both cases, the special color is grey.

@@

We now proceed to prove the above mentioned theorem.\\

@@

\noindent \underline{\textbf{First case:} \(h(x) \leqslant k\)} \\
\qquad Assume that \(h(x+y) \leqslant k\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of \(h\) 
and given that \(h(u) = h(v)\), \(f(u) = f(v)\). Distinguish between two cases according to the value of \(x + y\).

@@

\noindent \underline{\textbf{Second case:} \(h(x) \geqslant k + 1\)} \\
\qquad \(h(x) = k + g(\alpha) = 
k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
to the value of \(x + y\).

\begin{itemize}
\item If \(u + v > p\), write \(w = u + v - p \in [\![1, p]\!]\). Then \(x + y = (\alpha + \beta - 1) p + w\). Assume that 
	\(h(x+y) \geqslant k + 1\),  otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition, \(h(x + y) = k + g(\alpha + 
	\beta)\). Given that \(g(\alpha) = g(\beta)\), the sum-free property of \(g\) implies \(g(\alpha + \beta) 
	\neq g(\alpha)\) that is \(h(x + y) \neq h(x)\).
\item  If \(u + v \leqslant p\), write \(w = u + v \in [\![1, p]\!]\). Then \(x+y = (\alpha + \beta- 2) p + w\). The sum-free 
	property of \(f\) implies \(f(w) \neq f(u)\). Therefore \(f(w) \leqslant k\) and thus \(h(x + y) \leqslant k\). In particular,
	given that \(h(x) \geqslant k + 1\), \(h(x + y) \neq h(x)\).
\end{itemize}
\end{proof}

@@

\begin{proposition}
Let \((k, p) \in \mathbb{N}^*)^2\) and let \(f\) be a coloring associated to a SF-template with width \(p\) and \(k+1\) 
colors. Let \(b \in \mathbb{N}\) (\(b = m_{k+1} - 1\) works) and assume there there exists a coloring \(g\) of 
\([\![1, b]\!]\) with \(k+1\) colors such that:

@@

This proposition corresponds to the fact that sometimes a column is not the sum of two columns of a given color, but adding this 
column to the color would create sums in the color when applying the extension procedure. However, the last line does not 
intereact will all the columns when it comes to creating new sums. As a result, the hypotheses made on the coloring of the last 
row can be weakened.

We also have a similar construction theorem where only \(S^+\) is involved.

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with and \(p\) and \(k+1\) colors,
	and a \(SF\)-template with width \(q\) and \(n\) colors, then there exists \(SF\)-template with width \(pq\) and \((n+k)\).
\end{theorem}

And the associated inequality :

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S^+(n+k) \geqslant S^+(k+1)S^+(n) \]
\end{corollary}

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the SF property inherited 
from the second SF-template.
\end{proof}

@@

\begin{definition}
A SF-template with \(n\) colors is said to be symmetric if the partition in \(n\) sum-free subsets derived (with the additive constant) from this template is symmetric. 
A sum-free partition \(A_1, ..., A_n\) of \([\![1, p]\!]\) is said to be symmetric if for all \( x \in [\![1, p]\!]\), \(x\) and \(p + 1 - x\) belong to the same subset 
(except if \(x = p + 1 - x\)).
\end{definition}

Using a SAT solver, we exhibited SF-templates, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We have sought templates providing the greatest value of 
\((a, b)\) (for the lexicographic order). When the number of colors excceeded 5, in order to reduce the search space we 
looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix):
\[
\begin{array}{r r r r r}
	S(n + 1) & \geqslant & 3 S(n) & + & 1 \\
	S(n + 2) & \geqslant & 9 S(n) & + & 4 \\
	S(n + 3) & \geqslant & 33 S(n) & + & 6 \\
	S(n + 4) & \geqslant & 111 S(n) & + & 43 \\
	S(n + 5) & \geqslant & 380 S(n) & + & 148 \\
	S(n + 6) & \geqslant & 1\,140 S(n) & + & 528 \\
\end{array}
\]

@@

\par
The first 3 inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimal. 
Finally, the sixth one is obtained by combining (see below) the SF-template with width 380 and the one with width 3. 
Although we could not find a better SF-template with 7 colors, the last inequality is definetely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

@@

\begin{corollary}
The growth rate for Schur numbers (and Ramsey numbers \(R_n(3)\))  satisfies \(\gamma \geqslant \sqrt[5]{380} \approx 3.28\).
\end{corollary}

\begin{proof}[\textsc{Proof.}]
It is a mere consequence of the inequality \( S(n+5) \geqslant 380S(n) + 148\). As for Ramsey
numbers, the following inequality holds \(S(n) \leqslant R_n(3) - 2\) (see \cite{AbbottHanson}).
\end{proof}

@@

\par
In the next section, we provide similar results for weak Schur numbers. We introduce WSF-templates and a corresponding sequence, 
\(WS^+\). We then derive similar relations and construction method allowing us to find new lower bounds for weak Schur numbers.

>>>>

\begin{definition}
Let \((p,n) \in (\mathbb{N}^*)^2\). A SF-template with width \(p\) and \(n\) colors is defined as a partition of 
\([\![1,p]\!]\) into \(n\) sum-free subsets \(A_1, A_2, ..., A_n\) verifying
\[
\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
\Longrightarrow x+y-p \notin A_i
\]
\end{definition}

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty[\![\). We define \(S^+(n)\) as the maximal width of a SF-template with \(n\) colors. 
	\(S^+(n)\) is well defined and verifies
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with width \(p\) and \(k+1\) colors 
	and a partition of \([\![1,q]\!]\) into \(n\) sum-free subsets then there exists a partition of \([\![1,pq+m_{k+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{k+1}\) is the minimum of the special subset in the \(SF\)-template.
\end{theorem}

@@

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They vertically 
extend a sum-free partition by repeating it and they use an other sum-free partition to color the other half according 
to the line number. This way the "blocks" act as safe areas for each other. We give here an example for \(p = 9\), 
\(q = 4\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) (2 S(2) + 1) + S(2)\), both with Abbott and 
Hanson's construction and with a SF-template which is not included in Abbott and Hanson's construction. In both cases, 
the special color is grey.

@@

We now proceed to prove the above mentioned theorem.

@@

\noindent \underline{\textbf{First case:} \(h(x) \leqslant k\)}
\par
Assume that \(h(x+y) \leqslant k\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of \(h\) 
and given that \(h(u) = h(v)\), \(f(u) = f(v)\). Distinguish between two cases according to the value of \(x + y\).

@@

\noindent \underline{\textbf{Second case:} \(h(x) \geqslant k + 1\)}
\par
\(h(x) = k + g(\alpha) = 
k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
to the value of \(x + y\).

\begin{itemize}
\item \begin{sloppypar}
	If \(u + v > p\), write \(w = u + v - p \in [\![1, p]\!]\). Then \(x + y = (\alpha + \beta - 1) p + w\). Assume that 
	\({h(x+y) \geqslant k + 1}\),  otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition, \({h(x + y) = k + g(\alpha + 
	\beta)}\). Given that \(g(\alpha) = g(\beta)\), the sum-free property of \(g\) implies \(g(\alpha + \beta) 
	\neq g(\alpha)\) that is \(h(x + y) \neq h(x)\).
	\end{sloppypar}
\item  If \(u + v \leqslant p\), write \(w = u + v \in [\![1, p]\!]\). Then \(x+y = (\alpha + \beta- 2) p + w\). The sum-free 
	property of \(f\) implies \(f(w) \neq f(u)\). Therefore \(f(w) \leqslant k\) and thus \(h(x + y) \leqslant k\). In particular,
	given that \(h(x) \geqslant k + 1\), \(h(x + y) \neq h(x)\).
\end{itemize}
\end{proof}

@@

\begin{proposition}
Let \((p, k) \in \mathbb{N}^*)^2\) and let \(f\) be a coloring associated to a SF-template with width \(p\) and \(k+1\) 
colors. Let \(b \in \mathbb{N}\) (\(b = m_{k+1} - 1\) works) and assume there there exists a coloring \(g\) of 
\([\![1, b]\!]\) with \(k+1\) colors such that

@@

This proposition corresponds to the fact that sometimes a column is not the sum of two columns of a given color, but 
adding this column to the color would create sums in the color when applying the extension procedure. However, the 
last row does not intereact with the right-most columns when it comes to creating new sums. As a result, the hypotheses 
made on the coloring of the last row can be weakened.

There is a construction theorem for SF-templates as well.

\begin{theorem}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with width \(p\) and \(k+1\) colors,
and a SF-template with width \(q\) and \(n\) colors, then there exists \(SF\)-template with width \(pq\) and \((n+k)\) 
colors.
\end{theorem}

And the associated inequality

\begin{corollary}
Let \(n, k \in \mathbb{N}^*\). Then
\[ S^+(n+k) \geqslant S^+(k+1)S^+(n) \]
\end{corollary}

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the SF-template property inherited 
from the second SF-template.
\end{proof}

@@

\begin{definition}
A sum-free partition \(A_1, ..., A_n\) of \([\![1, p]\!]\) is said to be symmetric if for all \( x \in [\![1, p]\!]\), 
\(x\) and \(p + 1 - x\) belong to the same subset (except if \(x = p + 1 - x\)).

A SF-template with \(n\) colors is said to be symmetric if the partition into \(n\) sum-free subsets derived 
from this template by applying the extension procedure with a sum-free partition of length 1 is symmetric. 
\end{definition}

We exhibited SF-templates using a SAT solver, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We have sought templates providing the greatest value of 
\((a, b)\) (for the lexicographic order). When the number of colors excceeded 5, in order to reduce the search space we 
only looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix).
\[
\begin{array}{r r r r r}
	S(n + 1) & \geqslant & 3 S(n) & + & 1 \\
	S(n + 2) & \geqslant & 9 S(n) & + & 4 \\
	S(n + 3) & \geqslant & 33 S(n) & + & 6 \\
	S(n + 4) & \geqslant & 111 S(n) & + & 43 \\
	S(n + 5) & \geqslant & 380 S(n) & + & 148 \\
	S(n + 6) & \geqslant & 1\,140 S(n) & + & 528 \\
\end{array}
\]

@@

\par
The first 3 inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimum. 
Finally, the sixth one is obtained by combining a SF-template with width 380 and one with width 3. 
Although we could not find a better SF-template with 7 colors, the last inequality is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

@@

\begin{corollary}
\begin{sloppypar}
The growth rate for Schur numbers (and Ramsey numbers \(R_n(3)\))  satisfies \({\gamma \geqslant \sqrt[5]{380} \approx 3.28}\).
\end{sloppypar}
\end{corollary}

\begin{proof}[\textsc{Proof.}]
It is a mere consequence of the inequality \( S(n+5) \geqslant 380S(n) + 148\). As for Ramsey
numbers, the following inequality holds \(S(n) \leqslant R_n(3) - 2\) (see \cite{AbbottHanson}) hence the result.
\end{proof}

@@

\par
In the next section, we provide similar results for weak Schur numbers. We introduce WSF-templates and a corresponding sequence, 
\(\WS^+\). We then derive similar relations and a construction method allowing us to find new lower bounds for weak Schur numbers.
